---
title: "Equity Demographic and Transportability HRS Study"
author: "Emmanuel Fle Chea"
date: "`r Sys.Date()`"
output: html_document
---

# LIBRARIES & SETUP

```{r}
#############################################
# LIBRARIES & SETUP
#############################################

library(dplyr)
library(tidyr)
library(ggplot2)
library(survey)
library(pROC)
library(tableone)
library(knitr)
library(kableExtra)
library(gridExtra)
library(scales)
library(broom)
library(boot)
library(caret)
library(haven)
library(DiagrammeR)
library(DiagrammeRsvg)
library(rsvg)
library(cutpointr)
library(readr)
library(plotROC)
library(ROCR)

# Create output directories
dir.create("tables", showWarnings = FALSE)
dir.create("results", showWarnings = FALSE)
dir.create("figures", showWarnings = FALSE)
dir.create("figures/png", showWarnings = FALSE, recursive = TRUE)
dir.create("figures/tiff", showWarnings = FALSE, recursive = TRUE)

# Set ggplot theme
theme_set(
  theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 11),
      legend.position = "bottom"
    )
)

# Reproducibility
set.seed(2026)
```


# DATA LOADING & PREPARATION

# NOTE: These SAS files come from the HRS Restricted Data Agreement (RDA).
# Paths assume local decrypted storage per RDA rules.

```{r}
#############################################
# DATA LOADING & PREPARATION
#############################################

# Raw HRS files
trk2020        <- read_sas("trk2020tr_r.sas7bdat") # trk2020: HRS tracker file with demographics, weights, strata, PSU
neurobiomarker <- read_sas("neurobiomarker_sithara.sas7bdat") # neurobiomarker: plasma ATN biomarkers (Aβ42/40, pTau181, NfL, GFAP)
cogimp         <- read_sas("cogimp9220a_r.sas7bdat") # cogimp: cognitive impairment classification (Langa-Weir)

n_trk_initial <- nrow(trk2020)
n_bio_initial <- nrow(neurobiomarker)
n_cog_initial <- nrow(cogimp)

cat("Initial tracker N:", n_trk_initial, "\n")
cat("Initial biomarker N:", n_bio_initial, "\n")
cat("Initial cognition N:", n_cog_initial, "\n")

# Tracker data
trk2020_sub <- trk2020 %>%
  select(HHID, PN, GENDER, HISPANIC, RACE, PAGE, PVBSWGTR, SECU, STRATUM, SCHLYRS) %>%
  distinct() %>%
  mutate(HHID_PN = paste0(HHID, "_", PN))

# Biomarker data
neurobiomarker <- neurobiomarker %>%
  mutate(HHID_PN = paste0(HHID, "_", PN))
```



```{r}
# 2016 vascular conditions data
health2016 <- read_fwf(
  file = "H16C_R.da",
  fwf_positions(
    start = c(1, 7, 26, 28, 124),
    end   = c(6, 9, 26, 28, 124),
    col_names = c("HHID", "PN", "HYPERTENSION", "DIABETES", "STROKE")
  )
)

# In HRS, 1 = Yes, 5 = No for many health conditions; other codes = missing
health2016_clean <- health2016 %>%
  mutate(
    HYPERTENSION = ifelse(HYPERTENSION == 1, 1,
                     ifelse(HYPERTENSION == 5, 0, NA)),
    DIABETES = ifelse(DIABETES == 1, 1,
                 ifelse(DIABETES == 5, 0, NA)),
    STROKE = ifelse(STROKE == 1, 1,
               ifelse(STROKE == 5, 0, NA))
  )
```


# COGNITION RECONSTRUCTION (2016, 2018, 2020)

```{r}
#############################################
# COGNITION RECONSTRUCTION (2016, 2018, 2020)
#############################################

cog_long <- cogimp %>%
  mutate(
    HHID_PN = paste0(HHID, "_", PN),

    # 2016 (Wave 13)
    Cog_2016_IMRC  = R13IMRC,
    Cog_2016_DLRC  = R13DLRC,
    Cog_2016_SER7  = R13SER7,
    Cog_2016_BWC20 = R13BWC20,
    Cog_Score_2016 = Cog_2016_IMRC + Cog_2016_DLRC + Cog_2016_SER7 + Cog_2016_BWC20,

    # 2018 (Wave 14) proxy/self collapsed
    Cog_2018_IMRC  = coalesce(R14IMRCP,  R14IMRCW),
    Cog_2018_DLRC  = coalesce(R14DLRCP,  R14DLRCW),
    Cog_2018_SER7  = coalesce(R14SER7P,  R14SER7W),
    Cog_2018_BWC20 = coalesce(R14BWC20P, R14BWC20W),
    Cog_Score_2018 = Cog_2018_IMRC + Cog_2018_DLRC + Cog_2018_SER7 + Cog_2018_BWC20,

    # 2020 (Wave 15) proxy/self collapsed
    Cog_2020_IMRC  = coalesce(R15IMRCP,  R15IMRCW),
    Cog_2020_DLRC  = coalesce(R15DLRCP,  R15DLRCW),
    Cog_2020_SER7  = coalesce(R15SER7P,  R15SER7W),
    Cog_2020_BWC20 = coalesce(R15BWC20P, R15BWC20W),
    Cog_Score_2020 = Cog_2020_IMRC + Cog_2020_DLRC + Cog_2020_SER7 + Cog_2020_BWC20
  )

classify_cog <- function(score) {
  case_when(
    is.na(score) ~ NA_character_,
    score <= 6   ~ "Dementia",
    score <= 11  ~ "CIND",
    TRUE         ~ "Normal"
  )
}

cog_long <- cog_long %>%
  mutate(
    Dementia_IMP_2016 = classify_cog(Cog_Score_2016),
    Dementia_IMP_2018 = classify_cog(Cog_Score_2018),
    Dementia_IMP_2020 = classify_cog(Cog_Score_2020)
  )

# Subset cognition variables: raw items + summary scores
cogimp_sub <- cog_long %>%
  select(
    HHID_PN,
    # raw 2016 items
    R13IMRC, R13DLRC, R13SER7, R13BWC20,
    # raw 2018 items
    R14IMRCP, R14IMRCW, R14DLRCP, R14DLRCW,
    R14SER7P, R14SER7W, R14BWC20P, R14BWC20W,
    # raw 2020 items
    R15IMRCP, R15IMRCW, R15DLRCP, R15DLRCW,
    R15SER7P, R15SER7W, R15BWC20P, R15BWC20W,
    # summary scores
    Cog_Score_2016, Dementia_IMP_2016,
    Cog_Score_2018, Dementia_IMP_2018,
    Cog_Score_2020, Dementia_IMP_2020
  )

n_cog_complete <- nrow(cogimp_sub)
cat("Cognition subset N:", n_cog_complete, "\n")
```

# MERGE INTO FINAL ANALYTIC DATASET

```{r}
#############################################
# MERGE INTO FINAL ANALYTIC DATASET
#############################################

# Inner joins ensure analytic sample includes only participants with biomarkers, tracker data, and cognition. 
# Vascular data is optional, so left_join preserves all biomarker participants.

# Add merge key to health data
health2016_clean <- health2016_clean %>%
  mutate(HHID_PN = paste0(HHID, "_", PN))

# # Check for duplicate IDs before merging 
# stopifnot(anyDuplicated(neurobiomarker$HHID_PN) == 0) 
# stopifnot(anyDuplicated(trk2020_sub$HHID_PN) == 0) 
# stopifnot(anyDuplicated(cogimp_sub$HHID_PN) == 0)

# Merge all datasets
final <- neurobiomarker %>%
  inner_join(trk2020_sub, by = "HHID_PN") %>%
  inner_join(cogimp_sub, by = "HHID_PN") %>%
  left_join(health2016_clean, by = "HHID_PN")

n_merged <- nrow(final)
cat("Merged sample size:", n_merged, "\n")
```

# VASCULAR BURDEN VARIABLES

```{r}
#############################################
# VASCULAR BURDEN VARIABLES
#############################################

final <- final %>%
  mutate(
    CVD_Burden = HYPERTENSION + DIABETES + STROKE,
    CVD_Any = CVD_Burden > 0
  )

```


# ATN CLASSIFICATION

```{r}
#############################################
# ATN CLASSIFICATION
#############################################

# Thresholds based on published plasma biomarker literature: 
# A+: AB42/40 < 0.063 
# T+: pTau181 > 2.5 pg/mL 
# N+: NfL > 20 pg/mL or GFAP > 100 pg/mL

final <- final %>%
  mutate(
    # A: Amyloid (AB42_40_ratio)
    A_positive = case_when(
      is.na(AB42_40_ratio) ~ NA,
      AB42_40_ratio < 0.063 ~ TRUE,
      TRUE ~ FALSE
    ),

    # T: Tau (pTau181_recode)
    T_positive = case_when(
      is.na(pTau181_recode) ~ NA,
      pTau181_recode > 2.5 ~ TRUE,
      TRUE ~ FALSE
    ),

    # N: Neurodegeneration (NfL and/or GFAP)
    N_positive = case_when(
      is.na(NfL) | is.na(GFAP) ~ NA,
      NfL > 20 | GFAP > 100 ~ TRUE,
      TRUE ~ FALSE
    ),

    # ATN profile (8 categories)
    ATN_profile = case_when(
      is.na(A_positive) | is.na(T_positive) | is.na(N_positive) ~ NA_character_,
      !A_positive & !T_positive & !N_positive ~ "A-T-N-",
      A_positive & !T_positive & !N_positive ~ "A+T-N-",
      !A_positive & T_positive & !N_positive ~ "A-T+N-",
      !A_positive & !T_positive & N_positive ~ "A-T-N+",
      A_positive & T_positive & !N_positive ~ "A+T+N-",
      A_positive & !T_positive & N_positive ~ "A+T-N+",
      !A_positive & T_positive & N_positive ~ "A-T+N+",
      A_positive & T_positive & N_positive ~ "A+T+N+",
      TRUE ~ NA_character_
    ),

    # Simplified AD pathology classification
    AD_pathology = case_when(
      is.na(A_positive) | is.na(T_positive) ~ NA_character_,
      A_positive & T_positive ~ "AD pathology",
      A_positive & !T_positive ~ "Amyloid only",
      !A_positive & (T_positive | N_positive) ~ "Suspected non-AD",
      TRUE ~ "Normal biomarkers"
    )
  )

cat("ATN profile distribution:\n")
print(table(final$ATN_profile, useNA = "always"))
cat("\nAD pathology distribution:\n")
print(table(final$AD_pathology, useNA = "always"))
```



# DEMOGRAPHIC VARIABLES

```{r}
#############################################
# DEMOGRAPHIC VARIABLES
#############################################

# HISPANIC codes 1-3 represent Hispanic ethnicity; overrides race per standard practice

final <- final %>%
  mutate(
    Race_Ethnicity = case_when(
      HISPANIC %in% c(1, 2, 3) ~ "Hispanic",
      RACE == 1 ~ "White",
      RACE == 2 ~ "Black",
      RACE == 7 ~ "Other",
      TRUE ~ "Other"
    ),

    Sex = factor(GENDER, levels = c(1, 2), labels = c("Male", "Female")),

    Age_Group = cut(
      PAGE,
      breaks = c(0, 65, 75, 85, 120),
      labels = c("<65", "65-74", "75-84", "85+"),
      right = FALSE
    ),

    Education = case_when(
      is.na(SCHLYRS) ~ NA_character_,
      SCHLYRS < 12 ~ "Less than HS",
      SCHLYRS == 12 ~ "HS graduate",
      SCHLYRS %in% 13:15 ~ "Some college",
      SCHLYRS >= 16 ~ "College+"
    ),
    Education = factor(
      Education,
      levels = c("Less than HS", "HS graduate", "Some college", "College+")
    ),

    Low_Education = SCHLYRS < 12,

    Race_Sex_Group = case_when(
      Race_Ethnicity == "White" & Sex == "Female" ~ "White Women",
      Race_Ethnicity == "White" & Sex == "Male"   ~ "White Men",
      Race_Ethnicity == "Black" & Sex == "Female" ~ "Black Women",
      Race_Ethnicity == "Black" & Sex == "Male"   ~ "Black Men",
      TRUE ~ "Other"
    )
  )
```


# COGNITIVE VARIABLES

```{r}
#############################################
# COGNITIVE VARIABLES
#############################################

# A decline of ≥3 points over 4 years is commonly used to indicate meaningful cognitive decline

final <- final %>%
  mutate(
    Cog_Change = Cog_Score_2020 - Cog_Score_2016,
    Cog_Decline = case_when(
      is.na(Cog_Change) ~ NA,
      Cog_Change <= -3 ~ TRUE,
      TRUE ~ FALSE
    ),

    CogImp_2016 = Dementia_IMP_2016 %in% c("CIND", "Dementia"),

    Dementia_Progression = case_when(
      is.na(Dementia_IMP_2016) | is.na(Dementia_IMP_2020) ~ NA_character_,
      Dementia_IMP_2016 == "Normal" & Dementia_IMP_2020 == "Dementia" ~ "Progressed",
      Dementia_IMP_2016 == "CIND"   & Dementia_IMP_2020 == "Dementia" ~ "Progressed",
      Dementia_IMP_2016 == "Normal" & Dementia_IMP_2020 == "CIND"     ~ "Declined",
      TRUE ~ "Stable"
    )
  )
```


# MISSINGNESS VARIABLES

```{r}
#############################################
# MISSINGNESS VARIABLES
#############################################

# Complete case defined using baseline cognition (2016) + all biomarkers. 

final <- final %>%
  mutate(
    Missing_NfL        = is.na(NfL),
    Missing_GFAP       = is.na(GFAP),
    Missing_AB42_40    = is.na(AB42_40_ratio),
    Missing_pTau181    = is.na(pTau181_recode),
    Missing_Any_Biomarker = Missing_NfL | Missing_GFAP | Missing_AB42_40 | Missing_pTau181,
    Missing_Cog_2016   = is.na(Cog_Score_2016),
    Missing_Cog_2020   = is.na(Cog_Score_2020),
    Complete_Case      = !Missing_Any_Biomarker & !Missing_Cog_2016
  )

missingness_summary <- final %>%
  summarise(
    N_Total          = n(),
    N_Missing_NfL    = sum(Missing_NfL),
    N_Missing_GFAP   = sum(Missing_GFAP),
    N_Missing_AB     = sum(Missing_AB42_40),
    N_Missing_pTau   = sum(Missing_pTau181),
    N_Missing_Any_Bio= sum(Missing_Any_Biomarker),
    N_Complete       = sum(Complete_Case)
  )

write.csv(missingness_summary, "tables/missingness_summary.csv", row.names = FALSE)
print(missingness_summary)
```

# SURVEY DESIGN OBJECTS
# HRS uses a multistage area probability design with stratification and clustering

```{r}
#############################################
# SURVEY DESIGN OBJECTS (FIXED)
#############################################

# Remove rows with missing weights
# HRS requires sampling weights for valid population inference; cases with missing weights are excluded
final_weighted <- final %>% 
  filter(!is.na(PVBSWGTR))

svy_design <- svydesign(
  ids     = ~SECU,
  strata  = ~STRATUM,
  weights = ~PVBSWGTR,
  data    = final_weighted,
  nest    = TRUE
)

# Complete-case survey design for analyses requiring full biomarker + cognition data
svy_design_complete <- svydesign(
  ids     = ~SECU,
  strata  = ~STRATUM,
  weights = ~PVBSWGTR,
  data    = final_weighted %>% filter(Complete_Case),
  nest    = TRUE
)

cat("Survey design created successfully\n")
cat("Total sample with weights:", nrow(final_weighted), "\n")
cat("Complete cases with weights:", sum(final_weighted$Complete_Case), "\n")
```

# RACE DISTRIBUTION SUMMARIES

```{r}
#############################################
# RACE DISTRIBUTION SUMMARIES
#############################################

# Check for unexpected race categories
table(final$Race_Ethnicity, useNA = "ifany")

# Total N by Race
race_total <- final %>%
  group_by(Race_Ethnicity) %>%
  summarise(
    N_Total = n(),
    .groups = "drop"
  )

# Complete-case N by Race
race_complete <- final %>%
  group_by(Race_Ethnicity) %>%
  summarise(
    N_Complete = sum(Complete_Case, na.rm = TRUE),
    .groups = "drop"
  )

# Biomarker-available N by Race
race_biomarker <- final %>%
  group_by(Race_Ethnicity) %>%
  summarise(
    N_Biomarker_Available = sum(!Missing_Any_Biomarker, na.rm = TRUE),
    .groups = "drop"
  )

# Race × Sex N
# Intersectional counts used for fairness and subgroup analyses
race_sex_counts <- final %>%
  group_by(Race_Ethnicity, Sex) %>%
  summarise(
    N = n(),
    N_Complete = sum(Complete_Case, na.rm = TRUE),
    N_Biomarker_Available = sum(!Missing_Any_Biomarker, na.rm = TRUE),
    .groups = "drop"
  )

#############################################
# MERGE INTO ONE SUPPLEMENT TABLE
#############################################

# Merge race-level summaries; left_join preserves all race categories
race_summary <- race_total %>%
  left_join(race_complete, by = "Race_Ethnicity") %>%
  left_join(race_biomarker, by = "Race_Ethnicity") %>%
  arrange(desc(N_Total))

# Save tables
write.csv(race_summary, "tables/Table_Race_Summary.csv", row.names = FALSE)
write.csv(race_sex_counts, "tables/Table_Race_Sex_Summary.csv", row.names = FALSE)

# Display nicely
kable(race_summary, digits = 0,
      caption = "Supplement Table: Race Distribution, Complete Cases, and Biomarker Availability") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

kable(race_sex_counts, digits = 0,
      caption = "Supplement Table: Race × Sex Counts (Total, Complete Case, Biomarker Available)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```


# ANALYSIS 1: TRANSPORTABILITY FRAMEWORK

```{r}
#############################################
# ANALYSIS 1: TRANSPORTABILITY FRAMEWORK
#############################################

## 1A. ATN Prevalence - Weighted vs Unweighted

unweighted_prev <- final %>%
  filter(!is.na(ATN_profile)) %>%
  group_by(ATN_profile) %>%
  summarise(n = n(), .groups = "drop") %>%
  mutate(Unweighted_Pct = n / sum(n) * 100)

weighted_prev <- svytable(~ATN_profile, design = svy_design_complete) %>%
  prop.table() %>%
  as.data.frame()

colnames(weighted_prev) <- c("ATN_profile", "Weighted_Pct")

weighted_prev <- weighted_prev %>%
  mutate(Weighted_Pct = Weighted_Pct * 100)

transportability_atn <- unweighted_prev %>%
  left_join(weighted_prev, by = "ATN_profile") %>%
  mutate(
    Difference = Weighted_Pct - Unweighted_Pct,
    Ratio      = Weighted_Pct / Unweighted_Pct
  ) %>%
  arrange(desc(n))

## 1B. AD Pathology Prevalence - Weighted vs Unweighted

unweighted_ad <- final %>%
  filter(!is.na(AD_pathology)) %>%
  group_by(AD_pathology) %>%
  summarise(n = n(), .groups = "drop") %>%
  mutate(Unweighted_Pct = n / sum(n) * 100)

weighted_ad <- svytable(~AD_pathology, design = svy_design_complete) %>%
  prop.table() %>%
  as.data.frame()

colnames(weighted_ad) <- c("AD_pathology", "Weighted_Pct")

weighted_ad <- weighted_ad %>%
  mutate(Weighted_Pct = Weighted_Pct * 100)

transportability_ad <- unweighted_ad %>%
  left_join(weighted_ad, by = "AD_pathology") %>%
  mutate(
    Difference = Weighted_Pct - Unweighted_Pct,
    Ratio      = Weighted_Pct / Unweighted_Pct
  )

write.csv(transportability_ad, "tables/Table_Transportability_AD_Prevalence.csv", row.names = FALSE)

kable(transportability_ad, digits = 2,
      caption = "Table 2: AD Pathology Prevalence - Weighted vs Unweighted") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

## 1C. Biomarker-Cognition Associations - Weighted vs Unweighted

final_complete <- final %>% filter(Complete_Case)

unwt_model <- lm(
  Cog_Score_2016 ~ A_positive + T_positive + N_positive +
    PAGE + Sex + Race_Ethnicity + Education,
  data = final_complete
)

unwt_results <- tidy(unwt_model, conf.int = TRUE) %>%
  mutate(Model = "Unweighted")

wt_model <- svyglm(
  Cog_Score_2016 ~ A_positive + T_positive + N_positive +
    PAGE + Sex + Race_Ethnicity + Education,
  design = svy_design_complete
)

wt_results <- tidy(wt_model, conf.int = TRUE) %>%
  mutate(Model = "Weighted")

transportability_assoc <- bind_rows(unwt_results, wt_results) %>%
  filter(term %in% c("A_positiveTRUE", "T_positiveTRUE", "N_positiveTRUE")) %>%
  mutate(
    term = recode(
      term,
      "A_positiveTRUE" = "Amyloid+",
      "T_positiveTRUE" = "Tau+",
      "N_positiveTRUE" = "Neurodegeneration+"
    )
  ) %>%
  select(
    Model,
    Biomarker = term,
    Estimate  = estimate,
    SE        = std.error,
    CI_Lower  = conf.low,
    CI_Upper  = conf.high,
    P_Value   = p.value
  )

write.csv(transportability_assoc, "tables/Table_Transportability_Associations.csv", row.names = FALSE)

kable(transportability_assoc, digits = 3,
      caption = "Table 3: Biomarker-Cognition Associations - Weighted vs Unweighted") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

## 1D. Forest Plot - Weighted vs Unweighted Coefficients

plot_data <- transportability_assoc %>%
  mutate(Biomarker = factor(Biomarker,
                            levels = c("Amyloid+", "Tau+", "Neurodegeneration+")))

p_forest <- ggplot(plot_data, aes(x = Estimate, y = Biomarker, color = Model)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +

  # Updated error bars (horizontal)
  geom_errorbar(
    aes(xmin = CI_Lower, xmax = CI_Upper),
    width = 0.2,                      # replaces height
    position = position_dodge(width = 0.5),
    orientation = "y"                 # required for horizontal bars
  ) +

  geom_point(
    size = 4,
    position = position_dodge(width = 0.5)
  ) +

  scale_color_manual(values = c("Unweighted" = "#E69F00", "Weighted" = "#56B4E9")) +
  labs(
    title    = "Biomarker-Cognition Associations: Weighted vs Unweighted",
    subtitle = "Impact on cognitive score (TICSm) in 2016",
    x        = "Regression Coefficient (95% CI)",
    y        = "",
    color    = "Model"
  ) +
  theme(legend.position = "bottom")

print(p_forest)

ggsave("figures/png/Figure_Transportability_Forest.png", p_forest,
       width = 10, height = 6, dpi = 300)

ggsave("figures/tiff/Figure_Transportability_Forest.tiff", p_forest,
       width = 10, height = 6, dpi = 300, compression = "lzw")
```



```{r}
############################################################
# CLEAN OUTCOME VARIABLE
############################################################

# Remove the corrupted variable
final$CogImp_2016 <- NULL

# Recreate clean binary numeric outcome: 0 = Normal, 1 = CIND/Dementia
final$CogImp_2016 <- ifelse(final$Dementia_IMP_2016 == "Normal", 0, 1)
final$CogImp_2016 <- as.numeric(final$CogImp_2016)

# Factor version ONLY for global ROC
final$CogImp_2016_factor <- factor(
  final$CogImp_2016,
  levels = c(0, 1),
  labels = c("Normal", "Impaired")
)

# Verify
table(final$CogImp_2016, useNA = "ifany")
str(final$CogImp_2016)
```


```{r}
############################################################
# GLOBAL ROC (pROC)
############################################################

roc_obj <- pROC::roc(
  final$CogImp_2016_factor,
  final$NfL,
  quiet     = TRUE,
  direction = "<"
)

print(roc_obj)
pROC::ci.auc(roc_obj)
```


# ANALYSIS 2: INTERSECTIONAL SUBGROUPS (Race × Sex)

# Intersectional analyses were restricted to White and Black participants, the two largest racial groups in the 2016 HRS Venous Blood Study. Hispanic and Other racial groups were excluded from stratified analyses due to small sample sizes and high variance in biomarker distributions, which precluded stable subgroup estimation.

```{r}
#############################################
# ANALYSIS 2: INTERSECTIONAL SUBGROUPS (Race × Sex)
#############################################

# ---------------------------------------------------------
# Reusable Biomarker Plot Function (Trimming + Log + Jitter)
# ---------------------------------------------------------
plot_biomarkers <- function(data, group_var, biomarkers, log_scale = TRUE, trim = TRUE) {
  
  df <- data %>%
    dplyr::select(dplyr::all_of(group_var), dplyr::all_of(biomarkers)) %>%
    tidyr::pivot_longer(
      cols      = dplyr::all_of(biomarkers),
      names_to  = "Biomarker",
      values_to = "Value"
    ) %>%
    dplyr::filter(is.finite(Value))
  
  if (trim) {
    df <- df %>%
      dplyr::group_by(Biomarker) %>%
      dplyr::mutate(
        lower = stats::quantile(Value, 0.01, na.rm = TRUE),
        upper = stats::quantile(Value, 0.99, na.rm = TRUE)
      ) %>%
      dplyr::ungroup() %>%
      dplyr::filter(Value >= lower & Value <= upper)
  }
  
  if (log_scale) {
    df <- df %>% dplyr::mutate(Value = log10(Value))
  }
  
  df <- df %>%
    dplyr::mutate(
      Biomarker = dplyr::recode(
        Biomarker,
        "NfL"            = "Neurofilament Light (pg/mL)",
        "GFAP"           = "GFAP (pg/mL)",
        "pTau181_recode" = "pTau181 (pg/mL)"
      )
    )
  
  ggplot2::ggplot(df, ggplot2::aes(x = .data[[group_var]], y = Value, fill = .data[[group_var]])) +
    ggplot2::geom_violin(alpha = 0.6, trim = TRUE) +
    ggplot2::geom_boxplot(width = 0.15, outlier.alpha = 0) +
    ggplot2::geom_jitter(width = 0.15, alpha = 0.3, size = 1) +
    ggplot2::facet_wrap(~Biomarker, scales = "free_y", ncol = 1) +
    ggplot2::scale_fill_brewer(palette = "Set2") +
    ggplot2::labs(
      x   = "",
      y   = ifelse(log_scale, "Log10 Biomarker Level", "Biomarker Level"),
      fill = "Group"
    ) +
    ggplot2::theme(
      axis.text.x      = ggplot2::element_text(angle = 45, hjust = 1),
      legend.position  = "none"
    )
}

# ---------------------------------------------------------
# 2A. Sample Sizes by Intersectional Groups
# ---------------------------------------------------------
intersectional_n <- final %>%
  dplyr::filter(Race_Sex_Group != "Other") %>%
  dplyr::group_by(Race_Sex_Group) %>%
  dplyr::summarise(
    N_Total         = dplyr::n(),
    N_Complete      = sum(Complete_Case),
    N_ATN_Available = sum(!is.na(ATN_profile)),
    N_CogImp        = sum(CogImp_2016, na.rm = TRUE),  # numeric 0/1
    .groups         = "drop"
  )

write.csv(intersectional_n, "tables/Table_Intersectional_Sample_Sizes.csv", row.names = FALSE)

knitr::kable(
  intersectional_n,
  caption = "Table 4: Sample Sizes by Intersectional Groups"
) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))

# ---------------------------------------------------------
# 2B. ATN Prevalence by Race × Sex
# ---------------------------------------------------------
intersectional_atn <- final %>%
  dplyr::filter(Race_Sex_Group != "Other", !is.na(AD_pathology)) %>%
  dplyr::group_by(Race_Sex_Group, AD_pathology) %>%
  dplyr::summarise(n = dplyr::n(), .groups = "drop") %>%
  dplyr::group_by(Race_Sex_Group) %>%
  dplyr::mutate(
    Percent = n / sum(n) * 100,
    Total   = sum(n)
  ) %>%
  dplyr::ungroup()

write.csv(intersectional_atn, "tables/Table_Intersectional_ATN_Prevalence.csv", row.names = FALSE)

knitr::kable(
  intersectional_atn,
  digits  = 2,
  caption = "Table 5: ATN Prevalence by Race × Sex"
) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))

# ---------------------------------------------------------
# 2C. Biomarker Distributions by Race × Sex
# ---------------------------------------------------------
p_intersectional_bio <- plot_biomarkers(
  data      = final %>% dplyr::filter(Race_Sex_Group != "Other"),
  group_var = "Race_Sex_Group",
  biomarkers = c("NfL", "GFAP", "pTau181_recode"),
  log_scale = TRUE,
  trim      = TRUE
) +
  ggplot2::labs(
    title    = "Biomarker Distributions by Race × Sex",
    subtitle = "Intersectional analysis with log-scaling, trimming, and jittered points"
  )

print(p_intersectional_bio)

ggplot2::ggsave("figures/png/Figure_Intersectional_Biomarkers.png", p_intersectional_bio,
                width = 12, height = 10, dpi = 300)
ggplot2::ggsave("figures/tiff/Figure_Intersectional_Biomarkers.tiff", p_intersectional_bio,
                width = 12, height = 10, dpi = 300, compression = "lzw")

# ---------------------------------------------------------
# 2D. ROC Analysis by Race × Sex
# ---------------------------------------------------------
roc_results_list <- list()
subgroups <- c("White Women", "White Men", "Black Women", "Black Men")

for (group in subgroups) {
  
  group_data <- final %>%
    dplyr::filter(
      Race_Sex_Group == group,
      Complete_Case,
      !is.na(CogImp_2016),
      !is.na(NfL),
      !is.na(GFAP),
      !is.na(pTau181_recode)
    ) %>%
    dplyr::mutate(
      CogImp_num = CogImp_2016  
    )
  
  # Need at least two outcome classes
  if (length(unique(group_data$CogImp_num)) < 2) {
    message("Skipping subgroup ", group, ": only one outcome class.")
    next
  }
  
  run_roc <- function(marker, marker_name) {
    tryCatch({
      r  <- pROC::roc(group_data$CogImp_num, group_data[[marker]], quiet = TRUE, direction = "<")
      ci <- pROC::ci.auc(r)
      tibble::tibble(
        Group     = group,
        Biomarker = marker_name,
        N         = nrow(group_data),
        AUC       = as.numeric(pROC::auc(r)),
        AUC_Lower = ci[1],
        AUC_Upper = ci[3]
      )
    }, error = function(e) NULL)
  }
  
  roc_results_list[[group]] <- dplyr::bind_rows(
    run_roc("NfL", "NfL"),
    run_roc("GFAP", "GFAP"),
    run_roc("pTau181_recode", "pTau181")
  )
}

roc_results <- roc_results_list %>%
  purrr::discard(is.null) %>%
  dplyr::bind_rows()

if (nrow(roc_results) == 0) {
  warning("No valid ROC results were produced for any subgroup.")
} else {
  roc_results <- roc_results %>%
    dplyr::filter(
      is.finite(AUC),
      is.finite(AUC_Lower),
      is.finite(AUC_Upper)
    )
}

# Save results
write.csv(roc_results, "tables/Table_Intersectional_ROC.csv", row.names = FALSE)

# Display table only if non-empty
if (nrow(roc_results) > 0) {
  knitr::kable(
    roc_results,
    digits  = 3,
    caption = "Table 6: ROC Analysis by Race × Sex"
  ) %>%
    kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))
}

# ---------------------------------------------------------
# 2E. ROC Curves Plot
# ---------------------------------------------------------
if (nrow(roc_results) > 0) {
  
  p_roc <- ggplot2::ggplot(roc_results, ggplot2::aes(x = Group, y = AUC, color = Biomarker)) +
    ggplot2::geom_point(size = 4, position = ggplot2::position_dodge(width = 0.5)) +
    ggplot2::geom_errorbar(
      ggplot2::aes(ymin = AUC_Lower, ymax = AUC_Upper),
      width     = 0.2,
      position  = ggplot2::position_dodge(width = 0.5)
    ) +
    ggplot2::geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray50") +
    ggplot2::scale_color_brewer(palette = "Dark2") +
    ggplot2::labs(
      title    = "Biomarker Performance by Race × Sex",
      subtitle = "AUC for predicting cognitive impairment (CIND/Dementia)",
      caption  = "Note: Small sample sizes in some subgroups",
      x        = "",
      y        = "AUC (95% CI)",
      color    = "Biomarker"
    ) +
    ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1))
  
  print(p_roc)
  
  ggplot2::ggsave("figures/png/Figure_Intersectional_ROC.png", p_roc,
                  width = 12, height = 8, dpi = 300)
  ggplot2::ggsave("figures/tiff/Figure_Intersectional_ROC.tiff", p_roc,
                  width = 12, height = 8, dpi = 300, compression = "lzw")
}
```



# ANALYSIS 3: EDUCATION AS STRUCTURAL DISADVANTAGE

```{r}
#############################################
# ANALYSIS 3: EDUCATION AS STRUCTURAL DISADVANTAGE
#############################################

# ---------------------------------------------------------
# Helper: Clean biomarker data (remove non-finite + trim)
# ---------------------------------------------------------
clean_biomarkers <- function(df, value_var = "Value", trim = TRUE) {
  df <- df %>% filter(is.finite(.data[[value_var]]))

  if (trim) {
    df <- df %>%
      group_by(Biomarker) %>%
      mutate(
        lower = quantile(.data[[value_var]], 0.01, na.rm = TRUE),
        upper = quantile(.data[[value_var]], 0.99, na.rm = TRUE)
      ) %>%
      ungroup() %>%
      filter(.data[[value_var]] >= lower & .data[[value_var]] <= upper)
  }

  df
}

# ---------------------------------------------------------
# 3A. ATN Prevalence by Education
# ---------------------------------------------------------
edu_atn <- final %>%
  filter(!is.na(Education) & !is.na(AD_pathology)) %>%
  group_by(Education, AD_pathology) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(Education) %>%
  mutate(
    Percent = n / sum(n) * 100,
    Total   = sum(n)
  ) %>%
  ungroup()

write.csv(edu_atn, "tables/Table_Education_ATN_Prevalence.csv", row.names = FALSE)

kable(edu_atn, digits = 2,
      caption = "Table 7: ATN Prevalence by Education Level") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# ---------------------------------------------------------
# 3B. Biomarker Distributions by Education
# ---------------------------------------------------------
plot_data_edu <- final %>%
  filter(!is.na(Education)) %>%
  select(Education, NfL, GFAP, pTau181_recode, AB42_40_ratio) %>%
  pivot_longer(
    cols      = c(NfL, GFAP, pTau181_recode, AB42_40_ratio),
    names_to  = "Biomarker",
    values_to = "Value"
  ) %>%
  mutate(
    Biomarker = recode(
      Biomarker,
      "NfL"            = "Neurofilament Light",
      "GFAP"           = "GFAP",
      "pTau181_recode" = "pTau181",
      "AB42_40_ratio"  = "Aβ42/40 Ratio"
    )
  ) %>%
  clean_biomarkers(trim = TRUE)

p_edu_bio <- ggplot(plot_data_edu, aes(x = Education, y = Value, fill = Education)) +
  geom_violin(
    alpha = 0.6,
    quantiles = c(0.25, 0.5, 0.75)   
  ) +
  geom_boxplot(width = 0.15, outlier.alpha = 0) +
  facet_wrap(~Biomarker, scales = "free_y", ncol = 2) +
  scale_fill_brewer(palette = "RdYlBu") +
  labs(
    title    = "Biomarker Distributions by Education Level",
    subtitle = "Education as proxy for structural disadvantage",
    x        = "",
    y        = "Biomarker Level"
  ) +
  theme(
    axis.text.x     = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )

print(p_edu_bio)

ggsave("figures/png/Figure_Education_Biomarkers.png", p_edu_bio,
       width = 12, height = 10, dpi = 300)
ggsave("figures/tiff/Figure_Education_Biomarkers.tiff", p_edu_bio,
       width = 12, height = 10, dpi = 300, compression = "lzw")

# ---------------------------------------------------------
# 3C. Biomarker–Cognition Associations Stratified by Education
# ---------------------------------------------------------
edu_models_results <- list()

for (edu_level in levels(final$Education)) {
  edu_data <- final %>% filter(Education == edu_level & Complete_Case)

  if (nrow(edu_data) > 50) {
    model <- lm(
      Cog_Score_2016 ~ A_positive + T_positive + N_positive +
        PAGE + Sex + Race_Ethnicity,
      data = edu_data
    )

    results <- tidy(model, conf.int = TRUE) %>%
      filter(term %in% c("A_positiveTRUE", "T_positiveTRUE", "N_positiveTRUE")) %>%
      mutate(
        Education = edu_level,
        N         = nrow(edu_data),
        Biomarker = recode(
          term,
          "A_positiveTRUE" = "Amyloid+",
          "T_positiveTRUE" = "Tau+",
          "N_positiveTRUE" = "Neurodegeneration+"
        )
      ) %>%
      select(Education, N, Biomarker, estimate, std.error, conf.low, conf.high, p.value)

    edu_models_results[[edu_level]] <- results
  }
}

edu_stratified <- bind_rows(edu_models_results) %>%
  rename(
    Estimate = estimate,
    SE       = std.error,
    CI_Lower = conf.low,
    CI_Upper = conf.high,
    P_Value  = p.value
  )

write.csv(edu_stratified, "tables/Table_Education_Stratified_Associations.csv", row.names = FALSE)

kable(edu_stratified, digits = 3,
      caption = "Table 8: Biomarker–Cognition Associations Stratified by Education") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# ---------------------------------------------------------
# 3D. Education × Biomarker Interaction Models
# ---------------------------------------------------------
interaction_model <- lm(
  Cog_Score_2016 ~ (A_positive + T_positive + N_positive) * Low_Education +
    PAGE + Sex + Race_Ethnicity,
  data = final %>% filter(Complete_Case)
)

interaction_results <- tidy(interaction_model, conf.int = TRUE) %>%
  filter(grepl(":", term)) %>%
  mutate(
    Interaction = term %>%
      gsub("TRUE", "", .) %>%
      gsub("A_positive", "Amyloid+", .) %>%
      gsub("T_positive", "Tau+", .) %>%
      gsub("N_positive", "Neurodegeneration+", .) %>%
      gsub("Low_Education", " × Low Education", .)
  ) %>%
  select(
    Interaction,
    Estimate = estimate,
    SE       = std.error,
    CI_Lower = conf.low,
    CI_Upper = conf.high,
    P_Value  = p.value
  )

write.csv(interaction_results, "tables/Table_Education_Interactions.csv", row.names = FALSE)

kable(interaction_results, digits = 3,
      caption = "Table 9: Education × Biomarker Interactions") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# ---------------------------------------------------------
# 3E. Forest Plot — Education-Stratified Association
# ---------------------------------------------------------
p_edu_forest <- ggplot(
  edu_stratified,
  aes(x = Estimate, y = Education, color = Biomarker)
) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_errorbar(
    aes(xmin = CI_Lower, xmax = CI_Upper),
    width = 0.2,
    position = position_dodge(width = 0.6),
    orientation = "y"   # <-- FIXED: replaces deprecated geom_errorbarh()
  ) +
  geom_point(size = 4, position = position_dodge(width = 0.6)) +
  scale_color_brewer(palette = "Set1") +
  labs(
    title    = "Biomarker-Cognition Associations by Education Level",
    subtitle = "Differential associations reveal structural disadvantage",
    x        = "Regression Coefficient (95% CI)",
    y        = "",
    color    = "Biomarker"
  ) +
  theme(legend.position = "bottom")

print(p_edu_forest)

ggsave("figures/png/Figure_Education_Forest.png", p_edu_forest,
       width = 12, height = 8, dpi = 300)
ggsave("figures/tiff/Figure_Education_Forest.tiff", p_edu_forest,
       width = 12, height = 8, dpi = 300, compression = "lzw")
```


# ANALYSIS 4: FAIRNESS METRICS

```{r}
#############################################
# ANALYSIS 4: FAIRNESS METRICS
#############################################

## 4A. Define Prediction and Outcome

final <- final %>%
  mutate(
    Predicted_High_Risk = A_positive & T_positive,
    Actual_Outcome      = CogImp_2016
  )

## 4B. Fairness Metrics by Race

fairness_by_race <- final %>%
  filter(!is.na(Race_Ethnicity) & !is.na(Predicted_High_Risk) & !is.na(Actual_Outcome)) %>%
  group_by(Race_Ethnicity) %>%
  summarise(
    N  = n(),
    TP = sum(Predicted_High_Risk & Actual_Outcome),
    FP = sum(Predicted_High_Risk & !Actual_Outcome),
    TN = sum(!Predicted_High_Risk & !Actual_Outcome),
    FN = sum(!Predicted_High_Risk & Actual_Outcome),
    TPR = TP / (TP + FN),
    FPR = FP / (FP + TN),
    PPV = TP / (TP + FP),
    NPV = TN / (TN + FN),
    Accuracy = (TP + TN) / N,
    .groups  = "drop"
  ) %>%
  mutate(across(where(is.numeric) & !N, ~round(., 3)))

write.csv(fairness_by_race, "tables/Table_Fairness_Metrics_Race.csv", row.names = FALSE)

kable(fairness_by_race, digits = 3,
      caption = "Table 10: Fairness Metrics by Race/Ethnicity") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

## 4C. Fairness Metrics by Sex

fairness_by_sex <- final %>%
  filter(!is.na(Sex) & !is.na(Predicted_High_Risk) & !is.na(Actual_Outcome)) %>%
  group_by(Sex) %>%
  summarise(
    N  = n(),
    TP = sum(Predicted_High_Risk & Actual_Outcome),
    FP = sum(Predicted_High_Risk & !Actual_Outcome),
    TN = sum(!Predicted_High_Risk & !Actual_Outcome),
    FN = sum(!Predicted_High_Risk & Actual_Outcome),
    TPR = TP / (TP + FN),
    FPR = FP / (FP + TN),
    PPV = TP / (TP + FP),
    NPV = TN / (TN + FN),
    Accuracy = (TP + TN) / N,
    .groups  = "drop"
  ) %>%
  mutate(across(where(is.numeric) & !N, ~round(., 3)))

write.csv(fairness_by_sex, "tables/Table_Fairness_Metrics_Sex.csv", row.names = FALSE)

kable(fairness_by_sex, digits = 3,
      caption = "Table 11: Fairness Metrics by Sex") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

## 4D. Fairness Metrics by Intersectional Groups

fairness_intersectional <- final %>%
  filter(
    Race_Sex_Group != "Other" &
      !is.na(Predicted_High_Risk) &
      !is.na(Actual_Outcome)
  ) %>%
  group_by(Race_Sex_Group) %>%
  summarise(
    N  = n(),
    TP = sum(Predicted_High_Risk & Actual_Outcome),
    FP = sum(Predicted_High_Risk & !Actual_Outcome),
    TN = sum(!Predicted_High_Risk & !Actual_Outcome),
    FN = sum(!Predicted_High_Risk & Actual_Outcome),
    TPR = TP / (TP + FN),
    FPR = FP / (FP + TN),
    PPV = TP / (TP + FP),
    NPV = TN / (TN + FN),
    Accuracy = (TP + TN) / N,
    .groups  = "drop"
  ) %>%
  mutate(across(where(is.numeric) & !N, ~round(., 3)))

write.csv(fairness_intersectional, "tables/Table_Fairness_Metrics_Intersectional.csv", row.names = FALSE)

kable(fairness_intersectional, digits = 3,
      caption = "Table 12: Fairness Metrics by Race × Sex") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

## 4E. Disparities

white_tpr    <- fairness_by_race$TPR[fairness_by_race$Race_Ethnicity == "White"]
black_tpr    <- fairness_by_race$TPR[fairness_by_race$Race_Ethnicity == "Black"]
hispanic_tpr <- fairness_by_race$TPR[fairness_by_race$Race_Ethnicity == "Hispanic"]

disparities <- data.frame(
  Comparison = c("White - Black", "White - Hispanic", "Male - Female"),
  Metric     = "TPR",
  Disparity  = c(
    white_tpr - black_tpr,
    white_tpr - hispanic_tpr,
    fairness_by_sex$TPR[fairness_by_sex$Sex == "Male"] -
      fairness_by_sex$TPR[fairness_by_sex$Sex == "Female"]
  )
) %>%
  mutate(Disparity = round(Disparity, 3))

write.csv(disparities, "tables/Table_Fairness_Disparities.csv", row.names = FALSE)

kable(disparities, digits = 3,
      caption = "Table 13: Fairness Disparities (TPR Differences)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

## 4F. Fairness Metrics Visualization

fairness_plot_data <- bind_rows(
  fairness_by_race %>%
    select(Group = Race_Ethnicity, N, TPR, FPR, PPV, NPV) %>%
    mutate(Stratification = "Race/Ethnicity"),
  fairness_by_sex %>%
    mutate(Group = as.character(Sex)) %>%
    select(Group, N, TPR, FPR, PPV, NPV) %>%
    mutate(Stratification = "Sex")
) %>%
  pivot_longer(
    cols      = c(TPR, FPR, PPV, NPV),
    names_to  = "Metric",
    values_to = "Value"
  )

p_fairness <- ggplot(
  fairness_plot_data,
  aes(x = Group, y = Value, fill = Metric)
) +
  geom_col(position = position_dodge(width = 0.8), alpha = 0.8) +
  facet_wrap(~Stratification, scales = "free_x") +
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = percent_format()) +
  labs(
    title    = "Fairness Metrics Across Demographic Groups",
    subtitle = "TPR disparities indicate differential biomarker sensitivity",
    x        = "",
    y        = "Proportion",
    fill     = "Metric"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(p_fairness)

ggsave("figures/png/Figure_Fairness_Metrics.png", p_fairness,
       width = 12, height = 8, dpi = 300)
ggsave("figures/tiff/Figure_Fairness_Metrics.tiff", p_fairness,
       width = 12, height = 8, dpi = 300, compression = "lzw")
```


# ANALYSIS 5: MISSINGNESS & SENSITIVITY

```{r}
#############################################
# ANALYSIS 5: MISSINGNESS & SENSITIVITY
#############################################

# ---------------------------------------------------------
# 5A. Missingness by Demographics
# ---------------------------------------------------------
missingness_demo <- final %>%
  group_by(Race_Ethnicity, Sex) %>%
  summarise(
    N               = n(),
    Pct_Missing_NfL = mean(Missing_NfL, na.rm = TRUE) * 100,
    Pct_Missing_GFAP= mean(Missing_GFAP, na.rm = TRUE) * 100,
    Pct_Missing_AB  = mean(Missing_AB42_40, na.rm = TRUE) * 100,
    Pct_Missing_pTau= mean(Missing_pTau181, na.rm = TRUE) * 100,
    Pct_Missing_Any = mean(Missing_Any_Biomarker, na.rm = TRUE) * 100,
    .groups         = "drop"
  ) %>%
  arrange(desc(Pct_Missing_Any))

write.csv(missingness_demo, "tables/Table_Missingness_by_Demographics.csv", row.names = FALSE)

kable(missingness_demo, digits = 2,
      caption = "Table 14: Biomarker Missingness by Race/Ethnicity and Sex") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# ---------------------------------------------------------
# 5B. Missingness by Education
# ---------------------------------------------------------
missingness_edu <- final %>%
  filter(!is.na(Education)) %>%
  group_by(Education) %>%
  summarise(
    N               = n(),
    Pct_Missing_NfL = mean(Missing_NfL, na.rm = TRUE) * 100,
    Pct_Missing_GFAP= mean(Missing_GFAP, na.rm = TRUE) * 100,
    Pct_Missing_AB  = mean(Missing_AB42_40, na.rm = TRUE) * 100,
    Pct_Missing_pTau= mean(Missing_pTau181, na.rm = TRUE) * 100,
    Pct_Missing_Any = mean(Missing_Any_Biomarker, na.rm = TRUE) * 100,
    .groups         = "drop"
  )

write.csv(missingness_edu, "tables/Table_Missingness_by_Education.csv", row.names = FALSE)

kable(missingness_edu, digits = 2,
      caption = "Table 15: Biomarker Missingness by Education Level") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# ---------------------------------------------------------
# 5C. Predict Missingness (IPW for Complete Case)
# ---------------------------------------------------------

# Logistic regression for probability of being complete-case
miss_model <- glm(
  Complete_Case ~ PAGE + Sex + Race_Ethnicity + Education,
  data   = final,
  family = binomial()
)

# Predicted probability of being observed
final$miss_prob <- predict(miss_model, type = "response")

# Replace NA miss_prob with mean probability
final$miss_prob[is.na(final$miss_prob)] <- mean(final$miss_prob, na.rm = TRUE)

# Inverse probability weight
final$miss_weight <- ifelse(final$Complete_Case, 1 / final$miss_prob, 0)

# Replace NA weights with 0
final$miss_weight[is.na(final$miss_weight)] <- 0

# Combined weight = survey weight × missingness weight
final$combined_weight <- final$PVBSWGTR * final$miss_weight
final$combined_weight[is.na(final$combined_weight)] <- 0

# Truncate at 99th percentile
w99 <- quantile(final$combined_weight, 0.99, na.rm = TRUE)
final$combined_weight_trunc <- pmin(final$combined_weight, w99)

# Replace any remaining NA
final$combined_weight_trunc[is.na(final$combined_weight_trunc)] <- 0

cat("Missingness weights created\n")
cat("Mean weight:", mean(final$miss_weight[final$Complete_Case], na.rm = TRUE), "\n")
cat("Range:", range(final$miss_weight[final$Complete_Case], na.rm = TRUE), "\n")

# ---------------------------------------------------------
# 5D. Sensitivity Analysis - Missingness-Adjusted Survey Model
# ---------------------------------------------------------

# Ensure SECU and STRATUM are not missing
final_svy <- final %>%
  filter(Complete_Case, !is.na(SECU), !is.na(STRATUM))

svy_design_combined <- svydesign(
  ids     = ~SECU,
  strata  = ~STRATUM,
  weights = ~combined_weight_trunc,
  data    = final_svy,
  nest    = TRUE
)

sensitivity_model <- suppressWarnings(
  svyglm(
    Cog_Score_2016 ~ A_positive + T_positive + N_positive +
      PAGE + Sex + Race_Ethnicity + Education,
    design = svy_design_combined
  )
)

sensitivity_results <- tidy(sensitivity_model, conf.int = TRUE) %>%
  filter(term %in% c("A_positiveTRUE", "T_positiveTRUE", "N_positiveTRUE")) %>%
  mutate(
    Model = "Missingness-Adjusted",
    Biomarker = recode(
      term,
      "A_positiveTRUE" = "Amyloid+",
      "T_positiveTRUE" = "Tau+",
      "N_positiveTRUE" = "Neurodegeneration+"
    )
  ) %>%
  select(
    Model,
    Biomarker,
    Estimate  = estimate,
    SE        = std.error,
    CI_Lower  = conf.low,
    CI_Upper  = conf.high,
    P_Value   = p.value
  )

# Compare with original weighted model
comparison_sensitivity <- bind_rows(
  transportability_assoc %>% filter(Model == "Weighted"),
  sensitivity_results
)

write.csv(comparison_sensitivity, "tables/Table_Sensitivity_Missingness.csv", row.names = FALSE)

kable(comparison_sensitivity, digits = 3,
      caption = "Table 16: Sensitivity Analysis — Missingness-Adjusted Weights") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# ---------------------------------------------------------
# 5E. Missingness Visualization
# ---------------------------------------------------------
miss_plot_data <- missingness_demo %>%
  select(Race_Ethnicity, Sex, starts_with("Pct_")) %>%
  pivot_longer(
    cols      = starts_with("Pct_"),
    names_to  = "Biomarker",
    values_to = "Pct_Missing"
  ) %>%
  mutate(
    Biomarker = gsub("Pct_Missing_", "", Biomarker),
    Biomarker = recode(
      Biomarker,
      "NfL"  = "NfL",
      "GFAP" = "GFAP",
      "AB"   = "Aβ42/40",
      "pTau" = "pTau181",
      "Any"  = "Any Biomarker"
    ),
    Group = paste0(Race_Ethnicity, " - ", Sex)
  )

p_missingness <- ggplot(
  miss_plot_data,
  aes(x = Group, y = Pct_Missing, fill = Biomarker)
) +
  geom_col(position = position_dodge(width = 0.8)) +
  scale_fill_brewer(palette = "Set3") +
  labs(
    title    = "Biomarker Missingness by Demographics",
    subtitle = "Differential missingness may indicate systematic bias",
    x        = "",
    y        = "% Missing",
    fill     = "Biomarker"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(p_missingness)

ggsave("figures/png/Figure_Missingness.png", p_missingness,
       width = 12, height = 8, dpi = 300)
ggsave("figures/tiff/Figure_Missingness.tiff", p_missingness,
       width = 12, height = 8, dpi = 300, compression = "lzw")
```


# ANALYSIS 6: COMPREHENSIVE DEMOGRAPHICS TABLE

```{r}
#############################################
# ANALYSIS 6: COMPREHENSIVE DEMOGRAPHICS TABLE
#############################################

vars_table1 <- c(
  "PAGE", "Sex", "Race_Ethnicity", "Education", "SCHLYRS",
  "Cog_Score_2016", "Dementia_IMP_2016",
  "A_positive", "T_positive", "N_positive", "AD_pathology"
)

tab1_unwt <- CreateTableOne(
  vars       = vars_table1,
  data       = final %>% filter(Complete_Case),
  factorVars = c(
    "Sex", "Race_Ethnicity", "Education", "Dementia_IMP_2016",
    "A_positive", "T_positive", "N_positive", "AD_pathology"
  )
)

tab1_wt <- svyCreateTableOne(
  vars       = vars_table1,
  data       = svy_design_complete,
  factorVars = c(
    "Sex", "Race_Ethnicity", "Education", "Dementia_IMP_2016",
    "A_positive", "T_positive", "N_positive", "AD_pathology"
  )
)

tab1_unwt_df <- print(tab1_unwt, printToggle = FALSE)
tab1_wt_df   <- print(tab1_wt, printToggle = FALSE)

table1_combined <- cbind(tab1_unwt_df, tab1_wt_df)
colnames(table1_combined) <- c("Unweighted", "Weighted")

write.csv(table1_combined, "tables/Table1_Sample_Characteristics.csv")

kable(table1_combined,
      caption = "Table 1: Sample Characteristics (Weighted vs Unweighted)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```


# ANALYSIS 7: COGNITIVE GRADIENTS BY ATN

```{r}
#############################################
# ANALYSIS 7: COGNITIVE GRADIENTS BY ATN
#############################################

## 7A. Cognition by ATN Status

cog_by_atn <- final %>%
  filter(!is.na(ATN_profile) & !is.na(Cog_Score_2016)) %>%
  group_by(ATN_profile) %>%
  summarise(
    N          = n(),
    Mean_Cog   = mean(Cog_Score_2016, na.rm = TRUE),
    SD_Cog     = sd(Cog_Score_2016, na.rm = TRUE),
    Median_Cog = median(Cog_Score_2016, na.rm = TRUE),
    IQR_Cog    = IQR(Cog_Score_2016, na.rm = TRUE),
    .groups    = "drop"
  ) %>%
  arrange(desc(Mean_Cog))

write.csv(cog_by_atn, "tables/Table_Cognition_by_ATN.csv", row.names = FALSE)

kable(cog_by_atn, digits = 2,
      caption = "Table 17: Cognitive Performance by ATN Profile") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

## 7B. Cognition by ATN, Stratified by Race

cog_atn_race <- final %>%
  filter(!is.na(AD_pathology) & !is.na(Cog_Score_2016) & !is.na(Race_Ethnicity)) %>%
  group_by(Race_Ethnicity, AD_pathology) %>%
  summarise(
    N        = n(),
    Mean_Cog = mean(Cog_Score_2016, na.rm = TRUE),
    SD_Cog   = sd(Cog_Score_2016, na.rm = TRUE),
    .groups  = "drop"
  ) %>%
  pivot_wider(
    names_from  = AD_pathology,
    values_from = c(N, Mean_Cog, SD_Cog)
  )

write.csv(cog_atn_race, "tables/Table_Cognition_by_ATN_Race.csv", row.names = FALSE)

kable(cog_atn_race, digits = 2,
      caption = "Table 18: Cognitive Performance by ATN, Stratified by Race") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

## 7C. Cognitive Gradients Visualization

p_cog_gradients <- final %>%
  filter(!is.na(AD_pathology) & !is.na(Cog_Score_2016)) %>%
  ggplot(aes(x = AD_pathology, y = Cog_Score_2016, fill = AD_pathology)) +
  geom_violin(alpha = 0.6) +
  geom_boxplot(width = 0.2, outlier.alpha = 0.3) +
  facet_wrap(~Race_Ethnicity) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title    = "Cognitive Gradients by AD Pathology Status",
    subtitle = "Stratified by race/ethnicity",
    x        = "",
    y        = "Cognitive Score (TICSm)",
    fill     = "AD Pathology"
  ) +
  theme(
    axis.text.x     = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )

print(p_cog_gradients)

ggsave("figures/png/Figure_Cognitive_Gradients.png", p_cog_gradients,
       width = 12, height = 8, dpi = 300)
ggsave("figures/tiff/Figure_Cognitive_Gradients.tiff", p_cog_gradients,
       width = 12, height = 8, dpi = 300, compression = "lzw")
```


# ANALYSIS 8 - Race × Biomarker Interactions (Main + Supplementary)

```{r}
#############################################
# ANALYSIS 8: RACE × BIOMARKER INTERACTIONS
#############################################

# ---------------------------------------------------------
# 8A. Race × Biomarker Interaction Models
# ---------------------------------------------------------
race_interaction_model <- lm(
  Cog_Score_2016 ~ (A_positive + T_positive + N_positive) * Race_Ethnicity +
                   PAGE + Sex + Education,
  data = final %>% filter(Complete_Case)
)

race_interactions <- tidy(race_interaction_model, conf.int = TRUE) %>%
  filter(grepl(":", term) & grepl("Race_Ethnicity", term)) %>%
  mutate(
    term = gsub("TRUE", "", term),
    term = gsub("A_positive", "Amyloid+", term),
    term = gsub("T_positive", "Tau+", term),
    term = gsub("N_positive", "Neurodegeneration+", term)
  ) %>%
  select(
    Interaction = term,
    Estimate = estimate,
    SE = std.error,
    CI_Lower = conf.low,
    CI_Upper = conf.high,
    P_Value = p.value
  )

write.csv(race_interactions, "tables/Table_Race_Biomarker_Interactions.csv", row.names = FALSE)

kable(race_interactions, digits = 3,
      caption = "Table 19: Race × Biomarker Interactions") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# ---------------------------------------------------------
# 8B. MAIN FIGURE — Race × Amyloid Interaction
# ---------------------------------------------------------

pred_amyloid <- expand.grid(
  A_positive = c(TRUE, FALSE),
  T_positive = FALSE,
  N_positive = FALSE,
  Race_Ethnicity = c("White", "Black", "Hispanic"),
  PAGE = mean(final$PAGE, na.rm = TRUE),
  Sex = "Female",
  Education = "College+"
)

pred_amyloid$Predicted_Cog <- predict(race_interaction_model, newdata = pred_amyloid)

p_race_amyloid <- ggplot(pred_amyloid,
                         aes(x = Race_Ethnicity, y = Predicted_Cog,
                             fill = A_positive, group = A_positive)) +
  geom_col(position = position_dodge(width = 0.8), alpha = 0.7) +
  scale_fill_manual(values = c("FALSE" = "#4DAF4A", "TRUE" = "#E41A1C"),
                    labels = c("Amyloid-", "Amyloid+")) +
  labs(
    title = expression(bold("Race × Amyloid Interaction on Cognition")),
    subtitle = "Predicted cognitive scores (adjusted for age, sex, education)",
    x = "Race/Ethnicity",
    y = "Predicted Cognitive Score",
    fill = "Amyloid Status"
  ) +
  theme_minimal(base_size = 12)

print(p_race_amyloid)

ggsave("figures/png/Figure_Race_Amyloid_Interaction.png", p_race_amyloid,
       width = 12, height = 8, dpi = 300)
ggsave("figures/tiff/Figure_Race_Amyloid_Interaction.tiff", p_race_amyloid,
       width = 12, height = 8, dpi = 300, compression = "lzw")

# ---------------------------------------------------------
# 8C. SUPPLEMENTARY FIGURE — Race × Tau Interaction
# ---------------------------------------------------------

pred_tau <- expand.grid(
  A_positive = FALSE,
  T_positive = c(TRUE, FALSE),
  N_positive = FALSE,
  Race_Ethnicity = c("White", "Black", "Hispanic"),
  PAGE = mean(final$PAGE, na.rm = TRUE),
  Sex = "Female",
  Education = "College+"
)

pred_tau$Predicted_Cog <- predict(race_interaction_model, newdata = pred_tau)

p_race_tau <- ggplot(pred_tau,
                     aes(x = Race_Ethnicity, y = Predicted_Cog,
                         fill = T_positive, group = T_positive)) +
  geom_col(position = position_dodge(width = 0.8), alpha = 0.7) +
  scale_fill_manual(values = c("FALSE" = "#80B1D3", "TRUE" = "#377EB8"),
                    labels = c("Tau-", "Tau+")) +
  labs(
    title = expression(bold("Race × Tau Interaction on Cognition")),
    subtitle = "Predicted cognitive scores (adjusted for age, sex, education)",
    x = "Race/Ethnicity",
    y = "Predicted Cognitive Score",
    fill = "Tau Status"
  ) +
  theme_minimal(base_size = 12)

print(p_race_tau)

ggsave("figures/png/Figure_Race_Tau_Interaction.png", p_race_tau,
       width = 12, height = 8, dpi = 300)
ggsave("figures/tiff/Figure_Race_Tau_Interaction.tiff", p_race_tau,
       width = 12, height = 8, dpi = 300, compression = "lzw")

# ---------------------------------------------------------
# 8D. SUPPLEMENTARY FIGURE — Race × Neurodegeneration Interaction
# ---------------------------------------------------------

pred_neuro <- expand.grid(
  A_positive = FALSE,
  T_positive = FALSE,
  N_positive = c(TRUE, FALSE),
  Race_Ethnicity = c("White", "Black", "Hispanic"),
  PAGE = mean(final$PAGE, na.rm = TRUE),
  Sex = "Female",
  Education = "College+"
)

pred_neuro$Predicted_Cog <- predict(race_interaction_model, newdata = pred_neuro)

p_race_neuro <- ggplot(pred_neuro,
                       aes(x = Race_Ethnicity, y = Predicted_Cog,
                           fill = N_positive, group = N_positive)) +
  geom_col(position = position_dodge(width = 0.8), alpha = 0.7) +
  scale_fill_manual(values = c("FALSE" = "#A6D854", "TRUE" = "#66A61E"),
                    labels = c("Neurodegeneration-", "Neurodegeneration+")) +
  labs(
    title = expression(bold("Race × Neurodegeneration Interaction on Cognition")),
    subtitle = "Predicted cognitive scores (adjusted for age, sex, education)",
    x = "Race/Ethnicity",
    y = "Predicted Cognitive Score",
    fill = "N Status"
  ) +
  theme_minimal(base_size = 12)

print(p_race_neuro)

ggsave("figures/png/Figure_Race_N_Interaction.png", p_race_neuro,
       width = 12, height = 8, dpi = 300)
ggsave("figures/tiff/Figure_Race_N_Interaction.tiff", p_race_neuro,
       width = 12, height = 8, dpi = 300, compression = "lzw")
```


# ANALYSIS 9: CONCEPTUAL DIAGRAM

```{r}
#############################################
# ANALYSIS 9: CONCEPTUAL DIAGRAM
#############################################

conceptual_diagram <- grViz("
digraph conceptual_model {

  graph [
    layout = dot,
    rankdir = TB,
    fontname = Helvetica,
    bgcolor = white,

    # Bold title + extra spacing to prevent clipping
    label = <
      <B>Conceptual Framework Linking Social Determinants, Blood Biomarkers, and Cognitive Outcomes</B>
      <BR/><BR/>
    >,
    labelloc = t,
    fontsize = 24,

    # Increase top margin to avoid cutting off the title
    margin = 1.2
  ]

  node [shape = box, style = filled, fontname = Helvetica, fontsize = 12]

  race [label = 'Race/Ethnicity', fillcolor = '#E69F00', fontcolor = white]
  sex [label = 'Sex', fillcolor = '#E69F00', fontcolor = white]
  edu [label = 'Education\\n(Structural Disadvantage)', fillcolor = '#E69F00', fontcolor = white]
  age [label = 'Age', fillcolor = '#E69F00', fontcolor = white]

  amyloid [label = 'Amyloid\\n(Aβ42/40)', fillcolor = '#56B4E9', fontcolor = white]
  tau [label = 'Tau\\n(pTau181)', fillcolor = '#56B4E9', fontcolor = white]
  neuro [label = 'Neurodegeneration\\n(NfL, GFAP)', fillcolor = '#56B4E9', fontcolor = white]

  atn [label = 'ATN Profile', fillcolor = '#009E73', fontcolor = white]
  cog [label = 'Cognitive\\nPerformance', fillcolor = '#F0E442']

  fairness [label = 'Fairness & Equity', fillcolor = '#CC79A7', fontcolor = white, shape = ellipse]
  transport [label = 'Transportability', fillcolor = '#D55E00', fontcolor = white, shape = ellipse]

  race -> amyloid
  race -> tau
  race -> neuro

  sex -> amyloid
  sex -> tau
  sex -> neuro

  edu -> amyloid
  edu -> tau
  edu -> neuro

  age -> amyloid
  age -> tau
  age -> neuro

  amyloid -> atn
  tau -> atn
  neuro -> atn

  atn -> cog

  race -> cog [style = dashed]
  sex -> cog [style = dashed]
  edu -> cog [style = dashed]
  age -> cog [style = dashed]

  fairness -> amyloid [style = dotted, color = '#CC79A7']
  fairness -> tau [style = dotted, color = '#CC79A7']
  fairness -> neuro [style = dotted, color = '#CC79A7']
  fairness -> cog [style = dotted, color = '#CC79A7']

  transport -> atn [style = dotted, color = '#D55E00']
  transport -> cog [style = dotted, color = '#D55E00']

  {rank = same; race; sex; edu; age}
  {rank = same; amyloid; tau; neuro}
  {rank = same; fairness; transport}
}
")

# Convert to SVG
svg_obj <- DiagrammeRsvg::export_svg(conceptual_diagram)

# Save PNG
rsvg_png(charToRaw(svg_obj),
         file = "figures/png/Figure_Conceptual_Diagram.png",
         width = 4200, height = 3000)

# Save tif
bitmap <- rsvg(charToRaw(svg_obj), width = 4200, height = 3000)

ggsave("figures/tiff/Figure_Conceptual_Diagram.tiff",
     width = 4200, height = 3000, units = "px", compression = "lzw")
grid::grid.raster(bitmap)
dev.off()

print(conceptual_diagram)
```


# ANALYSIS 10: BOOTSTRAP CONFIDENCE INTERVALS (INTERNAL VALIDATION)

```{r}
################################################
# ANALYSIS 10: BOOTSTRAP CONFIDENCE INTERVALS (INTERNAL VALIDATION)
################################################

# -------------------------------------------------
# 10A. Bootstrap ATN Prevalence
# -------------------------------------------------
set.seed(123)

# Bootstrap function
boot_atn_prev <- function(data, indices) {
  d <- data[indices, ]
  prev <- table(d$ATN_profile) / nrow(d)
  return(prev)
}

# Run bootstrap
boot_results_atn <- boot(
  data = final %>% filter(!is.na(ATN_profile)),
  statistic = boot_atn_prev,
  R = 1000
)

# Extract confidence intervals
atn_levels <- levels(factor(final$ATN_profile))
boot_ci_atn <- data.frame(
  ATN_profile = atn_levels,
  Estimate = colMeans(boot_results_atn$t, na.rm = TRUE),
  CI_Lower = apply(boot_results_atn$t, 2, quantile, probs = 0.025, na.rm = TRUE),
  CI_Upper = apply(boot_results_atn$t, 2, quantile, probs = 0.975, na.rm = TRUE)
) %>%
  mutate(across(where(is.numeric), ~. * 100))

write.csv(boot_ci_atn, "tables/Table_Bootstrap_ATN_Prevalence.csv", row.names = FALSE)

kable(boot_ci_atn, digits = 2,
      caption = "Table 20: Bootstrap Confidence Intervals - ATN Prevalence") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# ------------------------------------------------
# 10B. Bootstrap Regression Coefficients
# ------------------------------------------------
set.seed(456)

# Bootstrap function for regression
boot_reg <- function(data, indices) {
  d <- data[indices, ]
  model <- lm(Cog_Score_2016 ~ A_positive + T_positive + N_positive + 
              PAGE + Sex + Race_Ethnicity + Education, data = d)
  return(coef(model)[c("A_positiveTRUE", "T_positiveTRUE", "N_positiveTRUE")])
}

# Run bootstrap
boot_results_reg <- boot(
  data = final %>% filter(Complete_Case),
  statistic = boot_reg,
  R = 1000
)

# Extract confidence intervals
boot_ci_reg <- data.frame(
  Biomarker = c("Amyloid+", "Tau+", "Neurodegeneration+"),
  Estimate = colMeans(boot_results_reg$t, na.rm = TRUE),
  CI_Lower = apply(boot_results_reg$t, 2, quantile, probs = 0.025, na.rm = TRUE),
  CI_Upper = apply(boot_results_reg$t, 2, quantile, probs = 0.975, na.rm = TRUE)
)

write.csv(boot_ci_reg, "tables/Table_Bootstrap_Regression_Coefficients.csv", row.names = FALSE)

kable(boot_ci_reg, digits = 3,
      caption = "Table 21: Bootstrap Confidence Intervals - Regression Coefficients") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```


# ANALYSIS 11: SENSITIVITY TO CUTPOINT DEFINITIONS

```{r}
########################################
# ANALYSIS 11: SENSITIVITY TO CUTPOINT DEFINITIONS
########################################

# ----------------------------------------------
# 11A. Cutpoint Sensitivity Analysis
# ----------------------------------------------
cutpoint_grid <- expand.grid(
  AB_cutpoint = c(0.060, 0.063, 0.065),
  pTau_cutpoint = c(2.0, 2.5, 3.0),
  NfL_cutpoint = c(15, 20, 25)
)

calc_prevalence <- function(AB_cut, pTau_cut, NfL_cut) {
  temp_data <- final %>%
    mutate(
      A_pos = AB42_40_ratio < AB_cut,
      T_pos = pTau181_recode > pTau_cut,
      N_pos = NfL > NfL_cut | GFAP > 100,
      AD_path = A_pos & T_pos
    ) %>%
    filter(!is.na(A_pos) & !is.na(T_pos) & !is.na(N_pos))
  
  data.frame(
    AB_cutpoint = AB_cut,
    pTau_cutpoint = pTau_cut,
    NfL_cutpoint = NfL_cut,
    Pct_A_pos = mean(temp_data$A_pos) * 100,
    Pct_T_pos = mean(temp_data$T_pos) * 100,
    Pct_N_pos = mean(temp_data$N_pos) * 100,
    Pct_AD_path = mean(temp_data$AD_path) * 100
  )
}

sensitivity_cutpoints <- cutpoint_grid %>%
  rowwise() %>%
  do(calc_prevalence(.$AB_cutpoint, .$pTau_cutpoint, .$NfL_cutpoint))

# ----------------------------------------------
# 11B. Heatmap Visualization
# ----------------------------------------------

sens_heatmap_data <- sensitivity_cutpoints %>%
  pivot_longer(
    cols = starts_with("Pct_"),
    names_to = "Outcome",
    values_to = "Percent"
  ) %>%
  mutate(
    Outcome = recode(Outcome,
                     "Pct_A_pos" = "Amyloid+",
                     "Pct_T_pos" = "Tau+",
                     "Pct_N_pos" = "Neurodegeneration+",
                     "Pct_AD_path" = "AD Pathology")
  )

p_cutpoint_heatmap <- ggplot(
  sens_heatmap_data,
  aes(x = factor(AB_cutpoint),
      y = factor(pTau_cutpoint),
      fill = Percent)
) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(option = "C", direction = -1) +
  facet_wrap(~Outcome + NfL_cutpoint, ncol = 3) +
  labs(
    title = expression(bold("Sensitivity to Cutpoint Definitions")),
    subtitle = "Heatmap of prevalence (%) across Aβ, pTau, and NfL thresholds",
    x = "Aβ Cutpoint",
    y = "pTau Cutpoint",
    fill = "Prevalence (%)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    strip.text = element_text(face = "bold"),
    panel.grid = element_blank()
  )

print(p_cutpoint_heatmap)

ggsave("figures/png/Figure_Cutpoint_Sensitivity.png", p_cutpoint_heatmap,
       width = 12, height = 8, dpi = 300)

ggsave("figures/tiff/Figure_Cutpoint_Sensitivity.tiff", p_cutpoint_heatmap,
       width = 12, height = 8, dpi = 300, compression = "lzw")
```


# ANALYSIS 12: COMPREHENSIVE RESULTS SUMMARY

```{r}
#######################################
# ANALYSIS 12: COMPREHENSIVE RESULTS SUMMARY
#######################################

# -----------------------------------------
# Create Results Summary Document
# -----------------------------------------
results_summary <- list(
  
  # Sample sizes
  Sample_Sizes = data.frame(
    Description = c("Initial biomarker sample", 
                   "After merging with tracker/cognition",
                   "Complete cases (all biomarkers + cognition)",
                   "Missing any biomarker",
                   "Missing cognition 2016"),
    N = c(n_bio_initial, n_merged, sum(final$Complete_Case),
          sum(final$Missing_Any_Biomarker), sum(final$Missing_Cog_2016))
  ),
  
  # ATN distribution
  ATN_Distribution = table(final$ATN_profile, useNA = "always") %>% 
    as.data.frame() %>%
    rename(ATN_Profile = Var1, N = Freq) %>%
    mutate(Percent = N / sum(N) * 100),
  
  # Demographics
  Demographics = final %>%
    filter(Complete_Case) %>%
    summarise(
      Mean_Age = mean(PAGE, na.rm = TRUE),
      SD_Age = sd(PAGE, na.rm = TRUE),
      Pct_Female = mean(Sex == "Female", na.rm = TRUE) * 100,
      Pct_White = mean(Race_Ethnicity == "White", na.rm = TRUE) * 100,
      Pct_Black = mean(Race_Ethnicity == "Black", na.rm = TRUE) * 100,
      Pct_Hispanic = mean(Race_Ethnicity == "Hispanic", na.rm = TRUE) * 100,
      Mean_Education = mean(SCHLYRS, na.rm = TRUE),
      Mean_Cognition = mean(Cog_Score_2016, na.rm = TRUE)
    ),
  
  # Key findings
  Key_Findings = data.frame(
    Finding = c(
      "Transportability: Weighted vs Unweighted ATN Prevalence Differs",
      "Education: Lower education associated with higher neurodegeneration",
      "Fairness: TPR disparities exist across race/ethnicity",
      "Intersectionality: Race × Sex interactions reveal differential patterns",
      "Missingness: Differential missingness by demographics"
    ),
    Evidence = c(
      "See Table_Transportability_ATN_Prevalence.csv",
      "See Table_Education_ATN_Prevalence.csv",
      "See Table_Fairness_Metrics_Race.csv",
      "See Table_Intersectional_ATN_Prevalence.csv",
      "See Table_Missingness_by_Demographics.csv"
    )
  )
)

# Save as RDS
saveRDS(results_summary, "results/results_summary.rds")

# Save as text
sink("results/results_summary.txt")

cat(strrep("=", 80), "\n")
cat("RESULTS SUMMARY: ATN EQUITY & TRANSPORTABILITY ANALYSIS\n")
cat(strrep("=", 80), "\n\n")

cat("SAMPLE SIZES:\n")
print(results_summary$Sample_Sizes)
cat("\n\n")

cat("ATN DISTRIBUTION:\n")
print(results_summary$ATN_Distribution)
cat("\n\n")

cat("DEMOGRAPHICS (COMPLETE CASES):\n")
print(results_summary$Demographics)
cat("\n\n")

cat("KEY FINDINGS:\n")
print(results_summary$Key_Findings)
cat("\n\n")

cat("Analysis completed:", Sys.time(), "\n")
sink()

cat("Results summary saved to results/results_summary.rds and results/results_summary.txt\n")
```


# ANALYSIS 13: VASCULAR COMORBIDITY BY RACE

```{r}
#############################################
# ANALYSIS 13: VASCULAR COMORBIDITY BY RACE
#############################################

vascular_by_race <- final %>%
  filter(Complete_Case) %>%
  group_by(Race_Ethnicity) %>%
  summarise(
    N = n(),
    Pct_Diabetes = mean(DIABETES, na.rm = TRUE) * 100,
    Pct_Hypertension = mean(HYPERTENSION, na.rm = TRUE) * 100,
    Pct_Stroke = mean(STROKE, na.rm = TRUE) * 100,
    Pct_Any_CVD = mean(CVD_Any, na.rm = TRUE) * 100,
    Mean_CVD_Burden = mean(CVD_Burden, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(Pct_Any_CVD))

write.csv(vascular_by_race, "tables/Table_S1_Vascular_Comorbidity_by_Race.csv", row.names = FALSE)

kable(vascular_by_race, digits = 1,
      caption = "Supplementary Table S1: Vascular Comorbidity Prevalence by Race/Ethnicity") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Vascular by Race × Sex
vascular_by_race_sex <- final %>%
  filter(Complete_Case, Race_Ethnicity %in% c("White", "Black")) %>%
  group_by(Race_Ethnicity, Sex) %>%
  summarise(
    N = n(),
    Pct_Diabetes = mean(DIABETES, na.rm = TRUE) * 100,
    Pct_Hypertension = mean(HYPERTENSION, na.rm = TRUE) * 100,
    Pct_Any_CVD = mean(CVD_Any, na.rm = TRUE) * 100,
    .groups = "drop"
  )

write.csv(vascular_by_race_sex, "tables/Table_S2_Vascular_by_Race_Sex.csv", row.names = FALSE)
```


# ANALYSIS 14: BIOMARKER-COGNITION ADJUSTED FOR VASCULAR BURDEN

```{r}
#############################################
# ANALYSIS 14: VASCULAR-ADJUSTED MODELS
#############################################

# Model 1: Base model (existing)
model_base <- svyglm(
  Cog_Score_2016 ~ A_positive + T_positive + N_positive +
    PAGE + Sex + Race_Ethnicity + Education,
  design = svy_design_complete
)

# Model 2: Add vascular comorbidities
model_vascular <- svyglm(
  Cog_Score_2016 ~ A_positive + T_positive + N_positive +
    PAGE + Sex + Race_Ethnicity + Education +
    DIABETES + HYPERTENSION + STROKE,
  design = svy_design_complete
)

# Compare models
base_results <- tidy(model_base, conf.int = TRUE) %>%
  filter(term %in% c("A_positiveTRUE", "T_positiveTRUE", "N_positiveTRUE")) %>%
  mutate(Model = "Base (Demographics Only)")

vascular_results <- tidy(model_vascular, conf.int = TRUE) %>%
  filter(term %in% c("A_positiveTRUE", "T_positiveTRUE", "N_positiveTRUE")) %>%
  mutate(Model = "Adjusted for Vascular Comorbidity")

comparison_vascular <- bind_rows(base_results, vascular_results) %>%
  mutate(
    Biomarker = recode(
      term,
      "A_positiveTRUE" = "Amyloid+",
      "T_positiveTRUE" = "Tau+",
      "N_positiveTRUE" = "Neurodegeneration+"
    )
  ) %>%
  select(Model, Biomarker, Estimate = estimate, SE = std.error,
         CI_Lower = conf.low, CI_Upper = conf.high, P_Value = p.value)

write.csv(comparison_vascular, "tables/Table_S3_Vascular_Adjusted_Associations.csv", row.names = FALSE)

kable(comparison_vascular, digits = 3,
      caption = "Supplementary Table S3: Biomarker-Cognition Associations With/Without Vascular Adjustment") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Forest plot comparing models
p_vascular_comparison <- ggplot(comparison_vascular,
                                aes(x = Estimate, y = Biomarker, color = Model)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_errorbar(
    aes(xmin = CI_Lower, xmax = CI_Upper),
    width = 0.2,
    position = position_dodge(width = 0.5),
    orientation = "y"
  ) +
  geom_point(size = 4, position = position_dodge(width = 0.5)) +
  scale_color_manual(values = c("Base (Demographics Only)" = "#E69F00",
                                 "Adjusted for Vascular Comorbidity" = "#56B4E9")) +
  labs(
    title = expression(bold("Impact of Vascular Comorbidity Adjustment on Biomarker-Cognition Associations")),
    subtitle = "Survey-weighted models",
    x = "Regression Coefficient (95% CI)",
    y = "",
    color = "Model"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")

print(p_vascular_comparison)

ggsave("figures/png/Figure_S1_Vascular_Adjustment_Comparison.png", p_vascular_comparison,
       width = 12, height = 8, dpi = 300)
ggsave("figures/tiff/Figure_S1_Vascular_Adjustment_Comparison.tiff", p_vascular_comparison,
       width = 12, height = 8, dpi = 300, compression = "lzw")
```


# ANALYSIS 15: CVD-STRATIFIED BIOMARKER ASSOCIATIONS

```{r}
#############################################
# ANALYSIS 15: BIOMARKER ASSOCIATIONS STRATIFIED BY CVD BURDEN
#############################################

# Stratify by CVD presence
cvd_strata_models <- list()

for (cvd_status in c("No CVD", "Any CVD")) {
  
  cvd_data <- if (cvd_status == "No CVD") {
    final %>% filter(Complete_Case, !CVD_Any)
  } else {
    final %>% filter(Complete_Case, CVD_Any)
  }
  
  if (nrow(cvd_data) > 100) {
    model <- lm(
      Cog_Score_2016 ~ A_positive + T_positive + N_positive +
        PAGE + Sex + Race_Ethnicity + Education,
      data = cvd_data
    )
    
    results <- tidy(model, conf.int = TRUE) %>%
      filter(term %in% c("A_positiveTRUE", "T_positiveTRUE", "N_positiveTRUE")) %>%
      mutate(
        CVD_Status = cvd_status,
        N = nrow(cvd_data),
        Biomarker = recode(
          term,
          "A_positiveTRUE" = "Amyloid+",
          "T_positiveTRUE" = "Tau+",
          "N_positiveTRUE" = "Neurodegeneration+"
        )
      )
    
    cvd_strata_models[[cvd_status]] <- results
  }
}

cvd_stratified_results <- bind_rows(cvd_strata_models) %>%
  select(CVD_Status, N, Biomarker, Estimate = estimate, SE = std.error,
         CI_Lower = conf.low, CI_Upper = conf.high, P_Value = p.value)

write.csv(cvd_stratified_results, "tables/Table_S4_CVD_Stratified_Associations.csv", row.names = FALSE)

kable(cvd_stratified_results, digits = 3,
      caption = "Supplementary Table S4: Biomarker-Cognition Associations Stratified by CVD Status") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Forest plot
p_cvd_stratified <- ggplot(cvd_stratified_results,
                           aes(x = Estimate, y = Biomarker, color = CVD_Status)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_errorbar(
    aes(xmin = CI_Lower, xmax = CI_Upper),
    width = 0.2,
    position = position_dodge(width = 0.5),
    orientation = "y"
  ) +
  geom_point(size = 4, position = position_dodge(width = 0.5)) +
  scale_color_manual(values = c("No CVD" = "#4DAF4A", "Any CVD" = "#E41A1C")) +
  labs(
    title = expression(bold("Biomarker-Cognition Associations Stratified by Vascular Comorbidity")),
    subtitle = "Testing whether CVD burden modulates biomarker effects",
    x = "Regression Coefficient (95% CI)",
    y = "",
    color = "CVD Status"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")

print(p_cvd_stratified)

ggsave("figures/png/Figure_S2_CVD_Stratified_Associations.png", p_cvd_stratified,
       width = 12, height = 8, dpi = 300)
ggsave("figures/tiff/Figure_S2_CVD_Stratified_Associations.tiff", p_cvd_stratified,
       width = 12, height = 8, dpi = 300, compression = "lzw")
```


# ANALYSIS 16: CALIBRATION PLOTS BY RACE

```{r}
#############################################
# ANALYSIS 16: BIOMARKER CALIBRATION BY RACE
#############################################

# Fit logistic regression for predicted risk
calibration_model <- glm(
  CogImp_2016 ~ A_positive + T_positive + N_positive,
  data = final %>% filter(Complete_Case),
  family = binomial()
)

final_calibration <- final %>%
  filter(Complete_Case) %>%
  mutate(
    Predicted_Risk = predict(calibration_model, type = "response"),
    Risk_Decile = ntile(Predicted_Risk, 10)
  )

# Calculate observed vs predicted by race and decile
calibration_data <- final_calibration %>%
  group_by(Race_Ethnicity, Risk_Decile) %>%
  summarise(
    N = n(),
    Predicted = mean(Predicted_Risk, na.rm = TRUE),
    Observed = mean(CogImp_2016, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(N >= 10)  # Minimum 10 per cell

write.csv(calibration_data, "tables/Table_S5_Calibration_Data.csv", row.names = FALSE)

# Calibration plot
p_calibration <- ggplot(calibration_data,
                       aes(x = Predicted, y = Observed, color = Race_Ethnicity)) +
  geom_point(aes(size = N), alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black") +
  geom_smooth(method = "loess", se = FALSE, linewidth = 1, formula = y ~ x) +
  scale_color_brewer(palette = "Dark2") +
  scale_size_continuous(range = c(2, 8)) +
  coord_cartesian(xlim = c(0, 0.5), ylim = c(0, 0.5)) +
  labs(
    title = expression(bold("Biomarker Calibration by Race/Ethnicity")),
    subtitle = "Predicted vs. Observed Cognitive Impairment Rates",
    caption = "Diagonal line = perfect calibration; deviation indicates under/over-prediction",
    x = "Predicted Risk of Cognitive Impairment",
    y = "Observed Rate of Cognitive Impairment",
    color = "Race/Ethnicity",
    size = "N per Decile"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right")

print(p_calibration)

ggsave("figures/png/Figure_S3_Calibration_by_Race.png", p_calibration,
       width = 12, height = 8, dpi = 300)
ggsave("figures/tiff/Figure_S3_Calibration_by_Race.tiff", p_calibration,
       width = 12, height = 8, dpi = 300, compression = "lzw")

# Calibration statistics
calibration_stats <- final_calibration %>%
  group_by(Race_Ethnicity) %>%
  summarise(
    N = n(),
    Mean_Predicted = mean(Predicted_Risk, na.rm = TRUE),
    Mean_Observed = mean(CogImp_2016, na.rm = TRUE),
    Calibration_Slope = coef(lm(CogImp_2016 ~ Predicted_Risk))[2],
    Calibration_Intercept = coef(lm(CogImp_2016 ~ Predicted_Risk))[1],
    .groups = "drop"
  )

write.csv(calibration_stats, "tables/Table_S6_Calibration_Statistics.csv", row.names = FALSE)

kable(calibration_stats, digits = 3,
      caption = "Supplementary Table S6: Calibration Statistics by Race/Ethnicity") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```


# ANALYSIS 17: YOUDEN INDEX OPTIMAL CUTPOINTS

```{r}
############################################################
# ANALYSIS 17: YOUDEN-OPTIMIZED CUTPOINTS + BOOTSTRAP CIs
############################################################

final <- final %>%
  mutate(
    CogImp_2016 = case_when(
      Dementia_IMP_2016 %in% c("CIND", "Dementia") ~ TRUE,
      Dementia_IMP_2016 == "Normal" ~ FALSE,
      TRUE ~ NA
    )
  )

############################################################
# FUNCTION 1 — Youden-optimal cutpoint
############################################################

optimize_cutpoint <- function(data, biomarker_var, outcome_var = "CogImp_2016") {

  data <- data %>%
    filter(
      !is.na(.data[[biomarker_var]]),
      !is.na(.data[[outcome_var]])
    )

  if (dplyr::n_distinct(data[[outcome_var]]) < 2) {
    return(data.frame(
      Biomarker        = biomarker_var,
      Optimal_Cutpoint = NA_real_,
      Youden_Index     = NA_real_,
      AUC              = NA_real_,
      Sensitivity      = NA_real_,
      Specificity      = NA_real_
    ))
  }

  res <- try(
    cutpointr(
      data   = data,
      x      = !!sym(biomarker_var),
      class  = !!sym(outcome_var),
      method = maximize_metric,
      metric = youden,
      silent = TRUE
    ),
    silent = TRUE
  )

  if (inherits(res, "try-error")) {
    return(data.frame(
      Biomarker        = biomarker_var,
      Optimal_Cutpoint = NA_real_,
      Youden_Index     = NA_real_,
      AUC              = NA_real_,
      Sensitivity      = NA_real_,
      Specificity      = NA_real_
    ))
  }

  data.frame(
    Biomarker        = biomarker_var,
    Optimal_Cutpoint = as.numeric(res$optimal_cutpoint),
    Youden_Index     = as.numeric(res$youden),
    AUC              = as.numeric(res$AUC),
    Sensitivity      = as.numeric(res$sensitivity),
    Specificity      = as.numeric(res$specificity)
  )
}

############################################################
# FUNCTION 2 — Bootstrap CI for cutpoints (robust, no NAs)
############################################################

bootstrap_cutpoint_ci <- function(data, biomarker_var,
                                  outcome_var = "CogImp_2016",
                                  B = 1000) {

  data <- data %>%
    filter(
      !is.na(.data[[biomarker_var]]),
      !is.na(.data[[outcome_var]])
    )

  if (dplyr::n_distinct(data[[outcome_var]]) < 2) {
    return(c(NA_real_, NA_real_))
  }

  boot_vals <- numeric(B)

  for (i in seq_len(B)) {

    samp <- data[sample(nrow(data), replace = TRUE), ]

    res <- try(
      cutpointr(
        data   = samp,
        x      = !!sym(biomarker_var),
        class  = !!sym(outcome_var),
        method = maximize_metric,
        metric = youden,
        silent = TRUE
      ),
      silent = TRUE
    )

    if (inherits(res, "try-error") || is.null(res$optimal_cutpoint)) {
      boot_vals[i] <- NA_real_
    } else {
      boot_vals[i] <- as.numeric(res$optimal_cutpoint)
    }
  }

  boot_vals <- boot_vals[!is.na(boot_vals)]

  if (length(boot_vals) == 0) {
    return(c(NA_real_, NA_real_))
  }

  if (length(unique(boot_vals)) == 1 || stats::sd(boot_vals) == 0) {
    m <- stats::median(boot_vals)
    return(c(m, m))
  }

  qs <- unname(stats::quantile(boot_vals, probs = c(0.025, 0.975), na.rm = TRUE))
  return(qs)
}

############################################################
# FUNCTION 3 — Add CI columns (FIXED)
############################################################

add_ci_columns <- function(df, data) {

  ci_results <- lapply(df$Biomarker, function(bio) {
    ci <- bootstrap_cutpoint_ci(data, bio)
    data.frame(CI_lower = ci[[1]], CI_upper = ci[[2]])
  })

  ci_df <- do.call(rbind, ci_results)

  df %>%
    bind_cols(ci_df)
}

############################################################
# OVERALL CUTPOINTS + CIs
############################################################

overall_cutpoints <- bind_rows(
  optimize_cutpoint(final, "AB42_40_ratio"),
  optimize_cutpoint(final, "pTau181_recode"),
  optimize_cutpoint(final, "NfL"),
  optimize_cutpoint(final, "GFAP")
) %>%
  mutate(Race = "Overall") %>%
  add_ci_columns(final)

############################################################
# RACE-SPECIFIC CUTPOINTS + CIs
############################################################

race_specific_cutpoints <- list()

for (race in c("White", "Black", "Hispanic")) {

  race_data <- final %>% filter(Race_Ethnicity == race)

  if (nrow(race_data) > 100) {

    race_cuts <- bind_rows(
      optimize_cutpoint(race_data, "AB42_40_ratio"),
      optimize_cutpoint(race_data, "pTau181_recode"),
      optimize_cutpoint(race_data, "NfL"),
      optimize_cutpoint(race_data, "GFAP")
    ) %>%
      mutate(Race = race) %>%
      add_ci_columns(race_data)

    race_specific_cutpoints[[race]] <- race_cuts
  }
}

race_cutpoints_df <- bind_rows(race_specific_cutpoints)

############################################################
# FINAL TABLE S7
############################################################

optimal_cutpoints_comparison <- bind_rows(
  overall_cutpoints,
  race_cutpoints_df
) %>%
  select(
    Race, Biomarker, Optimal_Cutpoint,
    CI_lower, CI_upper,
    Youden_Index, AUC, Sensitivity, Specificity
  )

write.csv(
  optimal_cutpoints_comparison,
  "tables/Table_S7_Optimal_Cutpoints_by_Race.csv",
  row.names = FALSE
)

kable(
  optimal_cutpoints_comparison,
  digits = 3,
  caption = "Supplementary Table S7: Youden-Optimized Biomarker Cutpoints by Race with 95% Bootstrap CIs"
) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

############################################################
# Visualization
############################################################

plot_overall <- overall_cutpoints %>% filter(!is.na(Optimal_Cutpoint))
plot_race    <- race_cutpoints_df %>% filter(!is.na(Optimal_Cutpoint))

if (nrow(plot_overall) == 0 || nrow(plot_race) == 0) {
  message("No valid cutpoints available for plotting.")
} else {

  p_cutpoint_comparison <- plot_race %>%
    ggplot(aes(x = Race, y = Optimal_Cutpoint, fill = Race)) +
    geom_col(alpha = 0.7) +
    geom_hline(
      data = plot_overall,
      aes(yintercept = Optimal_Cutpoint),
      linetype = "dashed",
      color = "red",
      linewidth = 1
    ) +
    facet_wrap(~ Biomarker, scales = "free_y", ncol = 2) +
    scale_fill_brewer(palette = "Set2") +
    labs(
      title = expression(bold("Race-Specific vs. Universal Optimal Biomarker Cutpoints")),
      subtitle = "Red dashed line = universal cutpoint; bars = race-specific Youden optima",
      y = "Optimal Cutpoint Value",
      x = ""
    ) +
    theme_minimal(base_size = 12) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "none"
    )

  print(p_cutpoint_comparison)

  ggsave(
    "figures/png/Figure_S4_Optimal_Cutpoints_Comparison.png",
    p_cutpoint_comparison,
    width = 12, height = 8, dpi = 300
  )

  ggsave(
    "figures/tiff/Figure_S4_Optimal_Cutpoints_Comparison.tiff",
    p_cutpoint_comparison,
    width = 12, height = 8, dpi = 300,
    compression = "lzw"
  )
}
```



# ANALYSIS 18: FAIRNESS LANDSCAPE HEATMAP

```{r}
#############################################
# ANALYSIS 18: FAIRNESS LANDSCAPE HEATMAP
#############################################

fairness_landscape <- final %>%
  filter(
    Race_Ethnicity %in% c("White", "Black"),
    !is.na(Education),
    !is.na(Predicted_High_Risk),
    !is.na(Actual_Outcome)
  ) %>%
  group_by(Race_Ethnicity, Sex, Education) %>%
  summarise(
    N = n(),
    TP = sum(Predicted_High_Risk & Actual_Outcome),
    FN = sum(!Predicted_High_Risk & Actual_Outcome),
    TPR = ifelse((TP + FN) > 0, TP / (TP + FN), NA),
    N_Impaired = sum(Actual_Outcome),
    .groups = "drop"
  ) %>%
  filter(N >= 20)  # Minimum cell size for stability

write.csv(fairness_landscape, "tables/Table_S8_Fairness_Landscape.csv", row.names = FALSE)

# Create heatmap
p_fairness_heatmap <- fairness_landscape %>%
  mutate(
    Group = interaction(Race_Ethnicity, Sex, sep = " "),
    Education = factor(Education, levels = c("Less than HS", "HS graduate", "Some college", "College+"))
  ) %>%
  ggplot(aes(x = Education, y = Group, fill = TPR)) +
  geom_tile(color = "white", linewidth = 1) +
  geom_text(aes(label = sprintf("%.1f%%\n(n=%d)", TPR*100, N)),
            color = "white", fontface = "bold", size = 3) +
  scale_fill_viridis_c(
    option = "plasma",
    limits = c(0, 0.35),
    labels = scales::percent,
    na.value = "gray80"
  ) +
  labs(
    title = expression(bold("Biomarker Sensitivity Landscape Across Intersectional Groups")),
    subtitle = "True Positive Rate (%) for detecting cognitive impairment via A+T+ biomarker profile",
    caption = "Cell values show TPR% and sample size. Higher TPR = better sensitivity.",
    x = "Educational Attainment",
    y = "Race × Sex Group",
    fill = "True Positive\nRate (TPR)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank(),
    legend.position = "right"
  )

print(p_fairness_heatmap)

ggsave("figures/png/Figure_Main_Fairness_Heatmap.png", p_fairness_heatmap,
       width = 12, height = 8, dpi = 300)
ggsave("figures/tiff/Figure_Main_Fairness_Heatmap.tiff", p_fairness_heatmap,
       width = 12, height = 8, dpi = 300, compression = "lzw")
```


# ANALYSIS 19: COHORT COMPARISON TABLE (ANALYSIS 19: HRS VS. ADNI COMPARISON)

```{r}
#############################################
# ANALYSIS 19: HRS VS. ADNI COMPARISON
#############################################

# Only variables with verified values are included.
# HRS values come from this study.
# ADNI values extracted from Weiner et al., 2013 (Alzheimers Dement).

cohort_comparison <- data.frame(
  Cohort   = c("HRS (This Study)", "ADNI (Weiner 2013)"),
  N        = c(4427, 416),
  Mean_Age = c(68.3, 75)   # ADNI mean age across AD/MCI/NC groups
)

write.csv(
  cohort_comparison,
  "tables/Table_S9_Cohort_Comparison.csv",
  row.names = FALSE
)

kable(
  cohort_comparison,
  digits = 1,
  caption = "Supplementary Table S9: HRS Compared to ADNI (Verified Variables Only)"
) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(1, background = "#FFF3CD")
```



# GENERATE COMPREHENSIVE REPORT

```{r}
##########################################
# GENERATE COMPREHENSIVE REPORT
#########################################

cat("\n")
cat(strrep("=", 80), "\n")
cat("ANALYSIS COMPLETE\n")
cat(strrep("=", 80), "\n\n")

cat("Tables generated:\n")
list.files("tables", pattern = "\\.csv$", full.names = FALSE) %>%
  paste("  -", .) %>%
  cat(sep = "\n")

cat("\n\nFigures generated (PNG):\n")
list.files("figures/png", pattern = "\\.png$", full.names = FALSE) %>%
  paste("  -", .) %>%
  cat(sep = "\n")

cat("\n\nFigures generated (tif):\n")
list.files("figures/tiff", pattern = "\\.tiff$", full.names = FALSE) %>%
  paste("  -", .) %>%
  cat(sep = "\n")

cat("\n\nResults files:\n")
list.files("results", full.names = FALSE) %>%
  paste("  -", .) %>%
  cat(sep = "\n")

cat("\n\n")
cat(strrep("=", 80), "\n")
cat("NEXT STEPS FOR HIGH-IMPACT PUBLICATION\n")
cat(strrep("=", 80), "\n\n")

cat("1. REVIEW ALL TABLES AND FIGURES\n")
cat("   - Check for accuracy and clarity\n")
cat("   - Ensure proper labeling and formatting\n\n")

cat("2. WRITE MANUSCRIPT SECTIONS\n")
cat("   - Introduction: Frame transportability and equity\n")
cat("   - Methods: Describe HRS, ATN classification, survey weighting\n")
cat("   - Results: Present analyses in logical order\n")
cat("   - Discussion: Emphasize population-level inference\n\n")

cat("3. KEY MESSAGES TO EMPHASIZE\n")
cat("   - HRS provides population-representative ATN data\n")
cat("   - Weighted analysis shows transportability\n")
cat("   - Education reflects structural disadvantage\n")
cat("   - Intersectional analysis reveals hidden disparities\n")
cat("   - Fairness metrics quantify equity gaps\n\n")

cat("4. TARGET JOURNALS (Ranked by Impact)\n")
cat("   - JAMA\n")
cat("   - JAMA Neurology\n")
cat("   - Alzheimer's & Dementia\n")
cat("   - Neurology\n")
cat("   - JAMA Network Open\n\n")

cat("5. NO EXTERNAL VALIDATION NEEDED\n")
cat("   - This is a descriptive equity study, not predictive modeling\n")
cat("   - HRS IS the population reference standard\n")
cat("   - Survey weights provide built-in transportability\n\n")

cat("Analysis timestamp:", format(Sys.time(), "%Y-%m-%d %H:%M:%S"), "\n")
cat(strrep("=", 80), "\n")

# SESSION INFO
cat("\nR Session Information:\n")
cat(strrep("=", 80), "\n")
sessionInfo()
sessionInfo()
```